INGRID: Creating Languages in MPS from ANTLR Grammars

### 2016-11-09 paper.tex
new latex class
small changes
abstract

\documentclass[10pt]{sigplanconf}

@@

% TODO
%\conferenceinfo{PLDI'17,} {June 19--26, 2017, Barcelona, Spain.}
%\CopyrightYear{2017}
%\copyrightdata{TODO}

@@

% TODO pridat nekoho dalsiho z JetBrains (Alex Shatalin?)
% TODO upravit affiliation pro JetBrains -> doplnit "s.r.o" nebo radsi "Inc."
% TODO muzeme taky napsat jejich emaily pokud budou lidi z JetBrains souhlasit
\authorinfo{V\'aclav Pech}
	{JetBrains}

@@

\begin{abstract}
% TODO
\todo{write later}

\todo{Copy from master thesis.}
JetBrains MPS is a language workbench focusing on domain-specific languages.
Unlike many other language workbenches and IDEs, it uses a projectional editor for code.
The developer directly manipulates the program in its tree form (AST) and not by editing a text source code.
This brings many advantages, but on the other hand requires time-consuming and complicated MPS language definition.
The thesis elaborates on the possibility of automating the process of creating MPS language definition from its grammar description.
It introduces the MPS editor, evaluates approaches of related projects and describes author's efforts to implement an MPS plugin that allows this import.
The chosen approach and the selection of tools used for implementation are justified in the thesis.
We point out important problems that any similar project might deal with and we introduce some possible solutions.
Furthermore, the thesis contains examples of imported languages, showing the potency of the chosen approach.
The thesis also aims to lay groundwork for future extensions and suggest possible improvements.
\end{abstract}

% TODO possible keywords: JetBrains MPS, programming language, grammar, projectional editor

>>>>

\documentclass[preprint,10pt,numbers]{sigplanconf}

@@

% TODO
%\conferenceinfo{PLDI '17}{June 19--26, 2017, Barcelona, Spain}
%\copyrightyear{20yy}
%\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
%\copyrightdoi{nnnnnnn.nnnnnnn}

% TODO
%\publicationrights{licensed}

@@

% TODO upravit affiliation pro JetBrains -> doplnit "s.r.o" nebo radsi "Inc."
% TODO muzeme taky napsat jejich emaily pokud budou lidi z JetBrains souhlasit
\authorinfo{V\'aclav Pech}
	{JetBrains}

@@

\begin{abstract}
JetBrains MPS is a language workbench, an IDE that allows developers both to write code and manipulate language definitions.
It leverages the concept of projectional editing, where the developer directly manipulates the AST representation of program code.
While MPS focuses on domain-specific languages (DSL), it needs to support also general-purpose programming languages (GPLs) --- for example, to compile and run programs written in some DSL, or to enable combination of GPLs with languages created in MPS.

We present \textsc{Ingrid}, a new method for construction of a language definition in MPS based on its ANTLR grammar.
The structure of a language is generated automatically, but its other aspects --- including the projectional editor --- must be adjusted manually by the user.
During the development of \textsc{Ingrid}, we encountered several challenges related to the mapping between (i) grammars written in an EBNF-like notation and (ii) the language definitions in MPS based on object-oriented principles.
Another difficulty is that grammars do not hold any information about code layout.

We implemented \textsc{Ingrid} as a plugin for MPS and evaluated it on several full-fledged programming languages, such as JavaScript and C\#.
Results show that our approach is practical.
Even though it is only partially automated, it saves the user from many hours of tedious and error-prone work.
Necessary manual adjustments take only a short time.
\end{abstract}

\category{D.3.2}{Language Classifications}{Extensible languages}
\category{D.3.4}{Processors}{Parsing}
\category{D.3.4}{Processors}{Code generation}
\category{D.2.3}{Coding Tools and Techniques}{Program editors}

\terms
Languages

\keywords
language workbench, projectional editor, grammar, domain-specific language, ANTLR, JetBrains MPS
